%!TEX root = generalexam.tex

% BEGIN ACKNOWLEDGEMENTS

This document was prepared in the span of 30 days in accordance to the rules put forth by the School of Meteorology's requirements for the General Examination. This work should be considered the thoughts and opinions of the author, and do not represent the opinions of his advisors.


Additionally, the author had very little knowledge of data assimilation when the project began. The first half of the allotted time was devoted to learning as much as possible about data assimilation. Everything from this time frame was, and still is, a blur. When it came time to write the report, it was very difficult to remember what source of information resulting in what knowledge. Thus, the author would like to acknowledge the following references that contributed to his understanding of data assimilation. This task would not have been able to be completed --- at least not with whatever sanity he has left --- without their willingness to make their papers, presentations, and lecture notes available online to the greater scientific community. In no particular order:

\vspace{2em}
\begin{itemize}
    \item \cite{evensen1994enkf} and \cite{evensen2003enkf}
    \vspace{1em}
    \item The ECMWF Lecture Series [Available online at: \url{http://www.ecmwf.int/newsevents/training/rcourse_notes}
    \vspace{1em}
    \item The power point presentations of Dr. Dale Barker from the Korean Meteorological Administration's Data Assimilation Workshops. Presentations can be found online at \url{http://weather.snu.ac.kr}
    \vspace{1em}
    \item A Presentation on Data Assimilation by Ms. Sarah L. Dance. [Available online at: \url{http://www.icfd.rdg.ac.uk/Workshops/Mathematics%20and%20Applications%20of%20Data%20Assimilation/S_Dance.pdf}]
    \vspace{1em}
\end{itemize}


\noindent Additionally, the author promised to convey that this document was written without the aid of caffeinated beverages (such as coffee, soda, or energy drinks). This document was written with the aid of lots of decaf iced-tea, many dozen homemade Reeses Pieces Peanut Butter cookies backed by his wife, and the sheer desire to complete his PhD.